---
title: "Demand_Analysis"
output:
  pdf_document: default
  html_document: default
date: "2024-04-28"
---

Install packages
```{r}
library(dplyr)
library(tidyr)
library(lavaan)
library(ggplot2)
library(scatterplot3d)
library(plotly)
```


Initializing the dataset
```{r}
# Read the data
data <- read.csv("raw_data/UpdatedSurveyApr29.csv", skip = 3, header = FALSE, na.strings = c("", " "), stringsAsFactors = FALSE)

# View the modified data
head(data)
summary(data)
```

Renaming the headers of the dataset for ease of use
```{r}
# Read the data
colnames(data) <- c("Familiarity", "WillingnessToTry", "Interest", "FlyingExperience", "AirshowExperience","PerceivedReliability", "EnvironmentalConcern", "WillingnessToPay", "ImportanceOfTimeSavings","ImportanceOfTrafficAvoidance", "ImportanceOfConvenience", "ImportanceOfEnvBenefits","ComfortWithFlying", "MainConcern", "DemonstrationInterest", "AAMUseCases", "OhioTesting","AgeGroup", "IncomeLevel", "AreaType")

# View the modified data
head(data)
summary(data)
#view(data)
#write.csv(data, "modified_data.csv", row.names = FALSE)
```

Converting Variables to Factors for Parsing
```{r}
# Convert WillingnessToTry, Interest, PerceivedReliability, EnvironmentalConcern, and ComfortWithFlying
factor_vars <- c("WillingnessToTry", "Interest", "PerceivedReliability", "EnvironmentalConcern", "ComfortWithFlying")
level_mapping <- c("Extremely unlikely" = -2, "Somewhat unlikely" = -1, "Neutral" = 0, "Somewhat likely" = 1, "Extremely likely" = 2,
                   "Extremely uninterested" = -2, "Somewhat uninterested" = -1, "Somewhat interested" = 1, "Extremely interested" = 2,
                   "Extremely unreliable" = -2, "Somewhat unreliable" = -1, "Somewhat reliable" = 1, "Extremely reliable" = 2,
                   "Not at all important" = -2, "Slightly important" = -1, "Moderately important" = 0, "Very important" = 1, "Extremely important" = 2,
                   "Extremely uncomfortable" = -2, "Somewhat uncomfortable" = -1, "Neither comfortable nor uncomfortable" = 0, "Somewhat comfortable" = 1, "Extremely comfortable" = 2)

for (var in factor_vars) {
  encoded_var <- paste0("enc_", var)
  data[[encoded_var]] <- factor(data[[var]], levels = names(level_mapping), labels = level_mapping)
}

# Convert FlyingExperience to binary (1 if any experience, 0 if none)
data$enc_FlyingExperience <- ifelse(data$FlyingExperience == "None of the above", 0, 1)

# Convert AirshowExperience, DemonstrationInterest, and OhioTesting
binary_vars <- c("Familiarity", "AirshowExperience", "DemonstrationInterest", "OhioTesting")
for (var in binary_vars) {
  encoded_var <- paste0("enc_", var)
  data[[encoded_var]] <- factor(data[[var]], levels = c("No", "Maybe", "Yes"), labels = c(-1, 0, 1))
}

# Convert WillingnessToPay
data$enc_WillingnessToPay <- factor(data$WillingnessToPay, levels = c("No more than a regular taxi/ride-share", "Up to 25% more", "Up to 50% more", "Up to 100% more (double)", "More than double"),
                                    labels = c(0, 0.25, 0.5, 1, 2))

# Convert IncomeLevel
data$enc_IncomeLevel <- factor(data$IncomeLevel, levels = c("Less than $25,000", "$25,000 - $50,000", "$50,001 - $75,000", "$75,001 - $100,000", "$100,001 - $150,000", "Over $150,000"),
                               labels = c(1, 2, 3, 4, 5, 6))

# Convert AreaType
data$enc_AreaType <- factor(data$AreaType, levels = c("Rural", "Suburban", "Urban"), labels = c(1, 2, 3))

# Save the modified dataset to a new CSV file
write.csv(data, "modified_data_encoded.csv", row.names = FALSE)
data <- read.csv("modified_data_encoded.csv")
summary(data)
```




Analyzing the Regression Correlation between demand factors: Interest in Trying + Willingness to Pay and Willingness to Try
```{r}
# Load the modified dataset with encoded values
data <- read.csv("modified_data_encoded.csv", stringsAsFactors = FALSE)

data_no_na <- na.omit(data)

model <- lm(enc_Interest ~ enc_WillingnessToPay + enc_WillingnessToTry, data = data_no_na)

# Extract the coefficients and R-squared
coefficients <- coef(model)
r_squared <- summary(model)$r.squared

# Print the coefficients and R-squared
print(coefficients)
print(r_squared)


# Residual plots
par(mfrow = c(2, 2))
plot(model, which = c(1, 2))

# Histogram of residuals
hist(model$residuals)

# QQ plot of residuals
qqnorm(model$residuals)
qqline(model$residuals)

# ANOVA table
anova(model)

# R-squared
summary(model)$r.squared

# Create the plot
ggplot(data = data_no_na, aes(x = enc_WillingnessToPay, y = enc_Interest)) +
  geom_point(color = "blue") +  # Add points
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add regression line
  labs(x = "Willingness to Pay More", y = "Interest in AAM") +  # Set labels
  theme_bw()  # Apply a theme (optional)

# Create the plot
ggplot(data = data_no_na, aes(x =enc_WillingnessToTry , y = enc_Interest)) +
  geom_point(color = "blue") +  # Add points
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add regression line
  labs(x = "Willingness to Try", y = "Interest in AAM") +  # Set labels
  theme_bw()  # Apply a theme (optional)

# Create the plot
ggplot(data = data_no_na, aes(x = enc_WillingnessToPay, y = enc_WillingnessToTry)) +
  geom_point(color = "blue") +  # Add points
  geom_smooth(method = "lm", se = TRUE, color = "red") +  # Add regression line
  labs(x = "Willingness to Pay", y = "Willingness to Try") +  # Set labels
  theme_bw()  # Apply a theme (optional)



library(plotly)


# Fit a linear regression model with all three variables
model_3d <- lm(enc_Interest ~ enc_WillingnessToPay + enc_WillingnessToTry, data = data_no_na)

#Analyzing the Regression Correlation for Public Acceptance
# Load the modified dataset with encoded values
data <- read.csv("modified_data_encoded.csv", stringsAsFactors = FALSE)

data_no_na <- na.omit(data)

model <- lm(enc_Interest ~ enc_WillingnessToTry, data = data_no_na)

# View the model summary
summary(model)
summary(model)$sigma #Standard Error of the Model
summary(model)$r.squared

# Extract the coefficients and R-squared
coefficients <- coef(model)
r_squared <- summary(model)$r.squared

# Print the coefficients and R-squared
print(coefficients)
print(r_squared)

```


Analyzing the Regression Correlation between acceptance factors: Interest in Trying + Willingness to Pay and Willingness to Try
```{r}
# Load the modified dataset with encoded values
data <- read.csv("modified_data_encoded.csv", stringsAsFactors = FALSE)
data_no_na <- na.omit(data)

# Fit the regression model
model <- lm(enc_ComfortWithFlying ~ enc_Familiarity + enc_PerceivedReliability, data = data_no_na)

# View the model summary
summary(model)

# Standard Error of the Model
summary(model)$sigma

# Residual plots
par(mfrow = c(2, 2))
plot(model, which = c(1, 2))

# Histogram of residuals
hist(model$residuals)

# QQ plot of residuals
qqnorm(model$residuals)
qqline(model$residuals)

# ANOVA table
anova(model)

# R-squared
summary(model)$r.squared

# Extract the coefficients and R-squared
coefficients <- coef(model)
r_squared <- summary(model)$r.squared

# Print the coefficients and R-squared
print(coefficients)
print(r_squared)

# Create scatter plots with regression lines
ggplot(data = data_no_na, aes(x = enc_Familiarity, y = enc_ComfortWithFlying)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", se = FLASE, color = "red") +
  labs(x = "Familiarity", y = "Comfort with Flying") +
  theme_bw()

ggplot(data = data_no_na, aes(x = enc_PerceivedReliability, y = enc_ComfortWithFlying)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "Perceived Reliability", y = "Comfort with Flying") +
  theme_bw()

# 3D scatter plot
library(plotly)

# Fit a linear regression model with all three variables
model_3d <- lm(enc_ComfortWithFlying ~ enc_Familiarity + enc_PerceivedReliability, data = data_no_na)

# Create a 3D scatter plot with the regression plane
fig <- plot_ly(data_no_na, x = ~enc_Familiarity, y = ~enc_PerceivedReliability, z = ~enc_ComfortWithFlying,
               marker = list(color = ~enc_ComfortWithFlying, colorscale = "Viridis"))

fig <- fig %>% add_markers()

fig <- fig %>% add_surface(x = ~enc_Familiarity, y = ~enc_PerceivedReliability, z = ~fitted(model_3d),
                           opacity = 0.5, colorscale = "Viridis")

fig <- fig %>% layout(scene = list(xaxis = list(title = "Familiarity"),
                                   yaxis = list(title = "Perceived Reliability"),
                                   zaxis = list(title = "Comfort with Flying")))


# Plot perceived reliability vs. familiarity
ggplot(data = data_no_na, aes(x = enc_Familiarity, y = enc_PerceivedReliability)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(x = "Familiarity", y = "Perceived Reliability") +
  theme_bw()

```
Creating the Demand & Acceptance Score
```{r}
# Load the modified dataset with encoded values
data <- read.csv("modified_data_encoded.csv", stringsAsFactors = FALSE)
data_no_na <- na.omit(data)

# Calculate Demand Score using PCA
demand_vars <- data_no_na[, c("enc_WillingnessToTry", "enc_WillingnessToPay", "enc_Interest")]
pca_demand <- prcomp(demand_vars, scale = TRUE)
data_no_na$DemandScore <- pca_demand$x[, 1]


# Calculate Acceptance Score using PCA
acceptance_vars <- data_no_na[, c("enc_Familiarity", "enc_PerceivedReliability", "enc_ComfortWithFlying")]
pca_acceptance <- prcomp(acceptance_vars, scale = TRUE)
data_no_na$AcceptanceScore <- pca_acceptance$x[, 1]

# Compare average scores
avg_demand_score <- mean(data_no_na$DemandScore)
avg_acceptance_score <- mean(data_no_na$AcceptanceScore)
print(paste("Average Demand Score:", avg_demand_score))
print(paste("Average Acceptance Score:", avg_acceptance_score))

# Calculate correlation between Demand Score and Acceptance Score
correlation <- cor(data_no_na$DemandScore, data_no_na$AcceptanceScore)
print(paste("Correlation between Demand Score and Acceptance Score:", correlation))

# Create a scatterplot
ggplot(data = data_no_na, aes(x = DemandScore, y = AcceptanceScore)) +
  geom_point(color = "blue") +
  labs(x = "Demand Score", y = "Acceptance Score") +
  theme_bw()

library(ggplot2)

ggplot(data_no_na, aes(x = DemandScore, y = AcceptanceScore)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "Demand Score", y = "Acceptance Score") +
  theme_minimal() +
  theme(plot.margin = margin(20, 20, 20, 20)) +
  geom_histogram(data = data_no_na, aes(x = DemandScore, y = ..density..), binwidth = 0.5,
                 fill = "blue", alpha = 0.6, position = "identity") +
  geom_histogram(data = data_no_na, aes(x = ..density.., y = AcceptanceScore), binwidth = 0.5,
                 fill = "green", alpha = 0.6, position = "identity") +
  coord_flip(xlim = c(-4, 4), ylim = c(-4, 4))

#Correlation Analysis
cor_coef <- cor(data_no_na$DemandScore, data_no_na$AcceptanceScore)
print(paste("Correlation Coefficient:", cor_coef))
#Output:
#[1] "Correlation Coefficient: 0.652989042712076"
#The correlation coefficient of 0.65 indicates a moderately strong positive correlation between the Demand Score and Acceptance Score. This suggests that as one score increases, the other tends to increase as well.

#Distribution Analysis 
par(mfrow = c(1, 2))
hist(data_no_na$DemandScore, main = "Distribution of Demand Score", xlab = "Demand Score")
hist(data_no_na$AcceptanceScore, main = "Distribution of Acceptance Score", xlab = "Acceptance Score")
#The histograms provide a visual representation of the distribution of each score. From the plots, we can observe that both scores appear to have a relatively normal distribution, with no extreme skewness or outliers.


#Mean and Variability
demand_mean <- mean(data_no_na$DemandScore)
demand_sd <- sd(data_no_na$DemandScore)
acceptance_mean <- mean(data_no_na$AcceptanceScore)
acceptance_sd <- sd(data_no_na$AcceptanceScore)

print(paste("Demand Score - Mean:", demand_mean, "| Standard Deviation:", demand_sd))
print(paste("Acceptance Score - Mean:", acceptance_mean, "| Standard Deviation:", acceptance_sd))
#Output:
#[1] "Demand Score - Mean: -0.143817453978833 | Standard Deviation: 1.41087252227541"
#[1] "Acceptance Score - Mean: -1.21932819031339 | Standard Deviation: 1.44841724882392"
#The Demand Score has a slightly higher mean (-0.14) compared to the Acceptance Score (-1.22), indicating a relatively higher level of demand. The standard deviations of both scores are similar, suggesting comparable variability.

#Practically, the specific goals and priorities of the AAM industry should guide the emphasis placed on each score. If the focus is on quick market penetration and capturing early adopters, the Demand Score may be more relevant. However, if the aim is to build trust and ensure long-term viability, the Acceptance Score may be more critical.
#Based on the analyses conducted, both the Demand Score and Acceptance Score appear to be valuable measures, with a moderately strong positive correlation between them. The choice of which score to prioritize depends on the specific objectives and strategies of the AAM industry.

```




Time Savings v. Traffic Avoidance v. Convenience v. Environmental Benefits
```{r}
# Calculate counts and percentages

data_long <- data %>%
  pivot_longer(cols = starts_with("ImportanceOf"),
               names_to = "category",
               values_to = "rank") 

top_ranked_summary <- data_long %>%
  filter(rank == 1) %>%
  na.omit() %>%
  group_by(category) %>%
  summarize(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  arrange(desc(percentage)) %>%
  mutate(category = recode(category, 
                           ImportanceOfConvenience = "Convenience",
                           ImportanceOfEnvBenefits = "Environmental Benefits",
                           ImportanceOfTimeSavings = "Time Savings",
                           ImportanceOfTrafficAvoidance = "Traffic Avoidance"))  # Rename categories

# Create bar chart with renamed x-axis labels
ggplot(top_ranked_summary, aes(x = reorder(category, -percentage), y = percentage)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "Factor", y = "Percentage of Respondents", 
       title = "Distribution of Top-Ranked Factors") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
summary(data)
```


Creating an Demand and Acceptance Score comparison with Demographics
```{r}
# Standardize relevant variables (excluding importance rankings)
data_scaled <- data %>%
  mutate(across(c(enc_Familiarity, enc_PerceivedReliability, enc_ComfortWithFlying, enc_WillingnessToTry, enc_WillingnessToPay, enc_Interest), scale))

# Remove rows with NAs before calculating scores
data_scaled <- na.omit(data_scaled)

# Calculate Demand Score using PCA
demand_vars <- data_scaled %>%
  select(enc_WillingnessToTry, enc_WillingnessToPay, enc_Interest)
pca_demand <- prcomp(demand_vars, scale = TRUE)
data_scaled$DemandScore <- pca_demand$x[, 1]

# Calculate Acceptance Score using PCA
acceptance_vars <- data_scaled %>%
  select(enc_Familiarity, enc_PerceivedReliability, enc_EnvironmentalConcern, enc_ComfortWithFlying)
pca_acceptance <- prcomp(acceptance_vars, scale = TRUE)
data_scaled$AcceptanceScore <- pca_acceptance$x[, 1]

# Standardize Demand and Acceptance Scores
data_scaled <- data_scaled %>%
  mutate(StandardizedDemandScore = scale(DemandScore),
         StandardizedAcceptanceScore = scale(AcceptanceScore))

# Plot Demand vs Acceptance with color-coding by age group
ggplot(data_scaled, aes(x = StandardizedDemandScore, y = StandardizedAcceptanceScore)) +
  geom_point(aes(color = AgeGroup), alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "Standardized Demand Score", y = "Standardized Acceptance Score",
       title = "Demand vs Acceptance for AAM") +
  scale_color_viridis_d() +
  theme_minimal()

# Calculate correlation between Demand and Acceptance
correlation <- cor(data_scaled$StandardizedDemandScore, data_scaled$StandardizedAcceptanceScore)
print(paste("Correlation between Demand and Acceptance:", correlation))

# Regression model to predict Acceptance based on Demand and other factors
model <- lm(StandardizedAcceptanceScore ~ StandardizedDemandScore + enc_IncomeLevel + enc_AreaType,
            data = data_scaled)
summary(model)

```

